Approximate State Abstraction

### 2016-01-21 paper/paper.tex
Some minor tweaks
quick skim of intro

% --- ABSTRACT ---
\begin{abstract}
The combinatorial explosion that plagues planning and \ac{RL} algorithms can be reversed using abstraction. For instance, prohibitively difficult task representations can be condensed so that solutions are tractably computable. In this work, we investigate a theoretical framework for approximate state abstraction that preserves near optimal behavior. \acp{RL} agents using these abstractions may treat experiences that resemble each other as equivalent, and generalize knowledge to novel scenarios based on prior experiences. We present theoretical guarantees of the quality of value functions derived from four classes of approximate state abstraction. Additionally, we empirically evaluate the relationship between the degree of approximation and the degree of abstraction achieved, as well as the tradeoff between approximation magnitude and optimality of behavior.
\end{abstract}

@@

% --- SECTION: Introduction ---
\section{Introduction}
Abstraction is a fundamental tool of intelligent agents seeking to reason about only the salient features of its environment. However, the knowledge available to an agent at any given moment is typically approximate and so agents are faced with the challenge of deciding when their knowledge is sufficiently accurate to form the basis of abstraction. This work focuses on characterizing what ``sufficiently accurate" information for abstraction amounts to in the context of planning and \ac{RL} in \acp{MDP}.

Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{littman1995complexity,papadimitriou1987complexity}. Similarly, many \ac{RL} algorithms for solving MDPs are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity is often what one hopes, due to Bellman's so-called curse of dimensionality the state space of an MDP often grows super-polynomially with the number of variables that characterize the domain. Thus, polynomial behavior in state space size is not necessarily the \enote{bee's knees}. For instance, a robot involved in a pick and place task might be able employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over CITE DABE ICAPS.

Thus, a key research interest of planning and RL has been leveraging abstraction as means of grappling with large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. These methods are characterized by reducing \textit{ground} MDPs with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states with similar values of interest. Existing work has characterized how aggregation of states with equal values of particular quantities fully maintain optimality~\cite{li2006towards,dean1997modelmin}. However, exactly solving for these quantities is itself computationally taxing and sometimes as difficult as solving for optimal behavior in the ground MDP, thereby defeating the purpose of abstraction.

In this work we demonstrate that aggregation of states on the basis of various approximately equal criteria incur bounded error.  Relaxing the aggregation criteria from equality of quantities to epsilon-closeness offers two benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge which we might reasonably a planning or learning algorithm to approximate without fully solving the MDP. Second, because the state aggregation criteria are relaxed to approximate equality, methods which employ approximate equality are able to tune the aggressiveness of state aggregation all the while incuring bounded error.

>>>>

% --- ABSTRACT ---
\begin{abstract}
The combinatorial explosion that plagues planning and \ac{RL} algorithms can be overcome using abstraction. For instance, prohibitively difficult task representations can be condensed so that solutions are tractably computable. In this work, we investigate a theoretical framework for approximate state abstraction that achieves high degrees of compressing while preserving near optimal behavior. \acp{RL} agents using these abstractions may treat experiences that resemble each other as equivalent, and generalize knowledge to novel scenarios based on prior experiences. We present theoretical guarantees of the quality of value functions derived from four classes of approximate state abstraction. Additionally, we empirically evaluate the relationship between the degree of approximation and the degree of abstraction achieved in a variety of tasks, as well as the tradeoff between approximation magnitude and optimality of behavior.
\end{abstract}

@@

% --- SECTION: Introduction ---
\section{Introduction}
Abstraction plays a fundamental role in learning; it encourages intelligent agents to reason about only the salient features of their environment while ignoring what is irrelevant, consequently enabling agents to solve considerably more complex problems. However, the knowledge available to a learning agent at any given moment is approximate. Thus, agents must determine when their knowledge is sufficiently accurate to form the basis of abstraction. This work characterizes what is ``sufficiently accurate" information for abstraction amounts to in the context of planning and \ac{RL} in \acp{MDP}.

Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{littman1995complexity,papadimitriou1987complexity}. Similarly, many \ac{RL} algorithms for solving MDPs are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, due to Bellman's so-called curse of dimensionality the size of the state space of an MDP grows super-polynomially with the number of variables that characterize the domain. Thus, polynomial behavior in state space size is not necessarily the \enote{bee's knees}. For instance, a robot involved in a pick and place task might be able employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over CITE DABE ICAPS.

Thus, a key research agenda for planning and RL has been leveraging abstraction as means of reducing large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. These methods are characterized by reducing \textit{ground} MDPs with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. Existing work has characterized how aggregation of states with equal values of particular quantities fully maintain optimality~\cite{li2006towards,dean1997modelmin}. However, exactly solving for these quantities is itself computationally taxing and sometimes as difficult as solving for optimal behavior in the ground MDP, thereby defeating the purpose of abstraction.

In this work we demonstrate that aggregation of states on the basis of various approximate criteria incur bounded error.  Relaxing the aggregation criteria from equality of quantities to $\varepsilon$-closeness offers two benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge which we expect a planning or learning algorithm to approximate without fully solving the MDP. Second, because the state aggregation criteria are relaxed to approximate equality, methods which employ approximate equality are able to tune the aggressiveness of state aggregation all the while incurring bounded error.
